A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time

### 2024-03-21 paper/main.tex
almost finished with results section, only full sample to write. next: discussion, conclusions, plot fixes, and intro

@@

%%% OUTLINE
\section{Methodology}
%recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
We utilize the group catalogs, produced by the \texttt{SUBFIND} algorithm~\citep{Springel2001,Dolag2009}, and the merger tree catalogs, generated by the \texttt{SUBLINK} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).

Our sample consists of major (stellar mass ratio $\M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated but physically associated.
From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of the subhalos of each pair to study their merger timescales.

@@

In addition to the physical separation, we also calculate the scaled separation of a pair at each snapshot. We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction of the virial radius $\rsep=X\Rvir$
% \begin{equation}
%     \rm \Rsc = \frac{\rsep}{\Rvir}
% \end{equation}
where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the primary's FoF group, which reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
This ``scaled radius" is, by construction, a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.

@@

Figure~\ref{fig:num-pairs} shows the number of low-mass and high-mass orbits as a function of redshift from $z=0-6$. 
Low-mass orbits (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
The number of pairs that merge decreases to 0 at $z=0$ since 
%pairs selected at $z=0$ by definition have not merged, and thus will not have the chance to, and 
many pairs that exist at low redshift will have merger timescales longer than the remaining time to the end of the simulation.

@@

\section{Results: The Mass and Redshift Dependence of Merger Timescales}
% The point of this subsection: to show that using the same "observability times" for pairs selected in the same physical separation bins wouldn't work, and that the selection criteria for close pairs should either be scaled with mass and redshift. 
\begin{figure*}[htb]
    \begin{center}
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
    \caption{\kc{Consider what redshifts to show. Either add colorbar normalized for all panels (in counts) or add \# of halos and a percentage. Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for low-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$. 
    There is a positive correlation between separation at the selected redshift and the time until merger. 
    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<100\,\kpc$, and will merge within $3-4\, \Gyr$. 
    Merger times are longer for lower redshift pairs, and the distribution of separations is larger.
    The lowest separation pairs have the shortest merger timescales.
    This plot can be used to get estimates for the merger timescale of an isolated low-mass pair at a given redshift. For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-6\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
    \kc{add z=5, z=6}
    }
    \label{fig:timevssep-low}
    \end{center}
\end{figure*}

\begin{figure*}[htb]
    \begin{center}
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
    \caption{\kc{Consider what redshifts to show. Add colorbar. Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for high-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
    Similar to the low-mass pairs, there is a positive correlation between the separation of a merging pair at a given redshift and the time until merger, with the lowest separation pairs tending to have the shortest time until merger.
    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<200\,\kpc$, and will merge within $0-7\, \Gyr$. 
    This plot can be used to get estimates for the merger timescale of an isolated high-mass pair at a given redshift. 
    For example, a high-mass pair at $z=2$ with $\rsep\sim 150\,\kpc$ will merge in $0.5-4\,\Gyr$, with a most likely time to merger of around $1.5-2\Gyr$.
    \kc{add z=5, z=6}
    }
    \label{fig:timevssep-high}
    \end{center}
\end{figure*}

\subsection{Time Until Merger vs. Separation}
% how we calculate for this section
We calculate the time until merger as a function of separation for low-mass and high-mass pairs at 6 different redshifts ($z=0.5,1,1.5,2,3,4$). 
At a given redshift, we identify all merging pairs that are post-infall and pre-merger, then find the associated time difference between the given redshift and the redshift of merger for each different pair.

% low mass example
For example, at $z=1.5$, there are XXX unique low-mass pairs that are post-infall and will merge prior to $z=0$, XXX (XX\%) of them have separations at $z=1.5$ between $\rsep=10-150\,kpc$, while XXX (XX\%) have $\rsep=10-50\,\kpc$. Of those with separations between $10-50\,\kpc$, the number of pairs with times until merger between $0-0.5\,\Gyr$, $0.5-1\,\Gyr$, $1-1.5\,\Gyr$, and $1.5-2\,\Gyr$ is $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, and $XX(XX\%)$ respectively.

% high mass example
Likewise, at $z=1.5$, there are XXX unique high-mass pairs that are post-infall and will merge prior to $z=0$, XXX (XX\%) of them have separations at $z=1.5$ between $\rsep=10-150\,kpc$, while XXX (XX\%) have $\rsep=10-50\,\kpc$. 
Of those with separations between $10-50\,\kpc$, the number of pairs with times until merger between $0-0.5\,\Gyr$, $0.5-1\,\Gyr$, $1-1.5\,\Gyr$, and $1.5-2\,\Gyr$ is $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, $XX(XX\%)$, and $XX(XX\%)$ respectively.
% might put the above in a table with more info?

% description of figures 
Figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of time until merger vs. separation for low-mass and high-mass pairs. 
In Fig.~\ref{fig:timevssep-low}, the time until merger for low-mass pairs is positively correlated to the separation of the pair at each redshift.

The figures can be used to get approximate `time until merger' ranges for a given separation at a given redshift.

% 
\input{table-septime.tex}

\subsection{Time Until Merger vs. Redshift}

Pairs chosen at $z=1.5$ on average merge at $z=XX$/within XX Gyrs.

\subsubsection{Separation-selected samples}

\subsection{"time til merger" for selected snapshots for LvH.}
Maybe also (distribution plot 2D hist? as function of selected separation)

% \begin{figure*}[htb]
%     \centering
%     \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_LoHi.png}
%     \caption{\kc{reverse the direction of z} 
%     The median (solid) and 1st-3rd quartile spread (shaded regions) of the cumulative time that isolated low-mass pairs (left) and high-mass pairs (right) spend with separations \rsep{} between $10\kpc$ and \Rphys{}. 
%     The Merging Pairs sample (dark green and dark pink) selected at $z=(1,1.5,2)$ merge before $z=0$, while the All Pairs sample (light green and light pink) include all the pairs from the \paircat{} (mergers and non-mergers).  
%     The low-mass (high-mass) merging pairs represent XX\%(XX\%) of the low-mass (high-mass) pair sample at $z=1,1.5, \mbox{and } 2$ respectively. 
%     The sample of All Pairs spend more time at higher separations than the merging pairs, as expected since the non-merging pairs are not likely to have very low separations that would result in mergers prior to $z=0$. 
%     For a more direct comparison between the low-mass and high-mass pairs, see Fig.~\ref{fig:phys-vs-scaled}. 
%     }
%     \label{fig:low-vs-high}
% \end{figure*}

% \begin{figure*}[htb]
%     \centering
%     \includegraphics[width=\textwidth]{plots/bet-on-it/2_medtotal_PhScT.png}
%     \caption{\kc{Add legend to plot, and I think reverse the direction of z} 
%     The median and 1st-3rd quartile spread of the cumulative time that isolated merging pairs, selected at $z=(1,1.5,2)$, spend with separations \rsep{} greater than $10\kpc$ and less than either \Rphys{} or \Rsc{}.
%     %
%     (Top) The median total time spent by isolated merging pairs with separations between $10\kpc$ and \Rphys{}.
%     Low-mass pairs spend a longer amount of time within $\sim200\kpc$ than high-mass pairs at each of these redshifts. 
%     However, high-mass galaxies spend more time at higher separations of $\rsep>200-300\kpc$. This is because the elapsed time is measured starting at the infall snapshot of the pair, and high-mass halos have larger radii, the high-mass halo orbits have a broader distribution of separations that extend to larger separations than those of the low-mass pairs.
%     (Bottom) The median total time spent by pairs with separations greater than $\rsep>10\kpc$ and less than a given fraction of the virial radius ($\rsep/\Rvir<\Rsc$) of the pair, which scales with mass and redshift.  
%     High-mass pairs spend more time within the same scaled separation than low-mass pairs. For example, a high-mass pair selected from the $z=1.5$ sample will spend \kc{XXGyr} within \kc{$1\Rsc\sim XX\kpc$}, while a low-mass pair from the same sample will spend \kc{XXGyr} within \kc{$1\Rsc\sim XX\kpc$}. 
%     Low-mass and high-mass pairs have similar cumulative time profiles as a function of scaled separation, with a roughly constant offset for all scaled separations. 
%     The high-mass pairs spend a median of \kc{$XX\Gyr$} more than low-mass pairs at every scaled separation at $z=1$ and $z=1.5$, and a median of \kc{$XX\Gyr$} more $z=2$.
%     }
%     \label{fig:phys-vs-scaled}
% \end{figure*}

@@

% \begin{figure*}[htb]
%     \centering
%     \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsz.png}
%     \caption{}
%     % \label{fig:low-vs-high}
% \end{figure*}
% \begin{figure*}[htb]
%     \centering
%     \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsz-2d.png}
%     \caption{}
%     % \label{fig:low-vs-high}
% \end{figure*}
\begin{figure*}[htb]
    \centering
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger_full_fit.png}
    \caption{\kc{TODO: add LookbackTime vs. Z line, add Hubble time line} The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs. Shaded regions represent the 1st and 3rd quartile spread on the median. Pairs at each redshift are selected if they have experienced first infall. 
    % 
    The black dashed line shows the Hubble time, which bounds the median time until merger from $z=6$ to $z\sim0.5-0.75$.
    The dotted grey line shows the time to $z=0$ (the Lookback Time) as a function of redshift, which sets the upper bound for the time until merger that a merging pair can have. 
    The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.}
    \label{fig:timescales}
\end{figure*}

@@

\section{Discussion}

@@

@@

\section{Summary and Conclusions}

% To do:
% -Sample: starting with FoF group cut and always show the total sample (merged and unmerged) since the unmerged sample is so small
% -Keep Fig 1: (two panels) add in a few more orbits to show the diversity and then have a panel to show one example where you shade everything below 100 kpc to show how the total time at S is computed
% -Keep Figure 2 as is 
% -Remove Figure 3
% -Figure 4: remove normalization of both samples so it's clear that the unmerged ones spend a lot of time at high separation and that they are a small percent of the fraction (put this in discussion or an appendix)
% -Remove Fig 5 and state conclusions in words along with the FoF group selection (most pairs have been in the same group for 1.5 Gyr) (can quote percentages from a cumulative version of the plot, or sum)
% -Figure 6: only show the *full* sample with the FoF group cut 
% check the half mass radius of the most massive dwarfs in your sample and justify 10 kpc by saying you want to be able to resolve twice that

% \section{Discussion}

% ** 
% \appendix
% adfadfa
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example1.png}
% \end{figure}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_example2.png}
%     \caption{}
% \end{figure}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/4_examplecombo.png}
% \end{figure}

% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_timeinbin_beforemerger.png}
% \end{figure}
% \begin{figure}[htb]
%     \centering
%     \includegraphics[width=0.5\columnwidth]{plots/bet-on-it/3_timeinbin_untilmerger_phys.png}
% \end{figure}

% \begin{figure*}[htb]
%     \centering
%     \includegraphics[width=\columnwidth]{plots/4_timescales/timevssep_phys+co.pdf}
%     \caption{}
%     % \label{fig:phys-comoving}
% \end{figure*}

% \begin{figure*}[htb]
%     \centering
%     \includegraphics[width=\columnwidth]{lotz.png}
%     \caption{}
%     % \label{fig:phys-comoving}
% \end{figure*}

>>>>

\newcommand{\chambe}{\citet{Chamberlain2024}}

@@

%%% OUTLINE
\section{Methodology} \label{sec:methods}
%recap last paper and mention which data sample we narrow down to (low mass major pairs only) -- give examples of orbits
We utilize the group catalogs, produced by the \texttt{SUBFIND} algorithm~\citep{Springel2001,Dolag2009}, and the merger tree catalogs, generated by the \texttt{SUBLINK} algorithm~\citep{RG2015}, from the highest resolution run of the IllustrisTNG simulation TNG100-1 (hereafter TNG100) to study the merger timescales of galaxy pairs from $z=0-6$.
The catalogs consist of a set of 100 snapshots, ranging from $z\sim20$ (snapshot 0) to $z=0$ (snapshot 99).

Our sample consists of major (stellar mass ratio $M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated but physically associated.
From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of the subhalos of each pair to study their merger timescales.

@@

In addition to the physical separation, we also calculate the scaled separation of a pair at each snapshot. We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction, $p$, of the virial radius $\rsep=p\Rvir$, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group.\footnote{Where the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.} 
The $\Rvir$ of the pair's FoF group reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
This ``scaled separation" is, by construction, a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.

@@

Figure~\ref{fig:numorbits} shows the number of low-mass and high-mass orbits as a function of redshift from $z=0-6$. 
Low-mass orbits (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
The number of pairs that merge decreases to 0 at $z=0$ since 
%pairs selected at $z=0$ by definition have not merged, and thus will not have the chance to, and 
many pairs that exist at low redshift will have merger timescales longer than the remaining time to the end of the simulation.

@@

\section{Results: The Mass and Redshift Dependence of Merger Timescales}
We have created catalogs of unique orbits for a large sample of isolated low-mass and high-mass pairs in the TNG100 simulation.
Next, we will calculate the merger timescales, or time until merger, for all of the merging orbits in our catalog. 
In Sec.~\ref{sec:results-timevsep}, we explore how the time until merger changes as a function of the separation of the pair at a variety of redshifts, and study how the time until merger changes between the low-mass sample and high-mass sample. 
In Sec.~\ref{sec:results-timevredshift}, we investigate the median time until merger for all post-infall and pre-merger isolated pairs as a function of redshift. 
We will additionally examine the merger timescale's dependence on a variety of separation criteria.

In the following analysis, we will utilize the orbit catalog, which consists of the unique orbits for all low-mass and high-mass major isolated pairs from the~\citet{Chamberlain2024} study. 
At a given redshift, we will include all merging pairs that are post-infall and pre-merger at that redshift in our analysis. 
However, separation criteria will also be applied to each redshift independently, so a pair that is in the $10-50\,\kpc$ bin at one redshift will not be part of the sample used at redshifts where it's separation is $>50\,\kpc$.

\begin{figure*}[htb]
    \begin{center}
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
    \caption{\kc{Add z=5, z=6 panels. Add colorbar for percentage per panel (and add number of pairs at that redshift) Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for low-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$. 
    There is a positive correlation between separation at the selected redshift and the time until merger. 
    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<100\,\kpc$, and will merge within $3-4\, \Gyr$. 
    Merger times are longer for lower redshift pairs, and the distribution of separations is larger.
    The lowest separation pairs have the shortest merger timescales.
    This plot can be used to get estimates for the merger timescale of an isolated low-mass pair at a given redshift. For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-6\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
    \kc{add z=5, z=6}
    }
    \label{fig:timevssep-low}
    \end{center}
\end{figure*}

\begin{figure*}[htb]
    \begin{center}
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
    \caption{\kc{Add z=5, z=6 panels. Add colorbar. Add low/high label somewhere, check the lower sep set to 10kpc?} The distribution of times until merger for high-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
    Similar to the low-mass pairs, there is a positive correlation between the separation of a merging pair at a given redshift and the time until merger, with the lowest separation pairs tending to have the shortest time until merger.
    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<200\,\kpc$, and will merge within $0-7\, \Gyr$. 
    This plot can be used to get estimates for the merger timescale of an isolated high-mass pair at a given redshift. 
    For example, a high-mass pair at $z=2$ with $\rsep\sim 150\,\kpc$ will merge in $0.5-4\,\Gyr$, with a most likely time to merger of around $1.5-2\Gyr$.
    }
    \label{fig:timevssep-high}
    \end{center}
\end{figure*}

\subsection{Separation Dependence of Merger Timescales}\label{sec:results-timevsep}
\kc{note: this might go in the appendix as additional info}
% how we calculate for this section
We calculate the time until merger as a function of separation for low-mass and high-mass pairs at a variety of redshifts from $z=0.5-6$. 
At the given redshift, we identify all merging pairs that are post-infall and pre-merger, then find the associated time difference between the given redshift and the redshift of merger for each pair.

% low mass example
For example, at $z=1.5$, there are 5,787 unique orbits that are post-infall and will merge prior to $z=0$. Of those, and 2,458 (42.47\%) have low separations between $\rsep=10-50\,\kpc$, and 2,207 of those low-separation pairs will merge in the next $2\,\Gyr$. 
A compilation of this data is available in Table~\ref{tab:timevssepvsz}, where we list the fraction of post-infall and pre-merger pairs within 3 separation bins $(10-50, 10-70, 10-100)\,\kpc$, and the fraction within each of those separation bins that merge in 4 different timescale bins $(0-1, 1-2, 2-3, 3+)\,\Gyr$, for redshifts of $z=(0.5,1,1.5,2,3,4,5,6)$..

As an accompanying visual aid, figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of time until merger vs. separation for low-mass and high-mass pairs at redshifts $z=(0.5,1,1.5,2,3,4,5,6)$. 
Fig.~\ref{fig:timevssep-low} presents the distribution for low-mass subhalos, and shows that the time until merger for low-mass pairs is positively correlated with the separation of the pair at each redshift, but that the positive correlation is smaller at larger redshifts. 
In addition, the spread of times until merger and separations is smaller at higher redshift  than the spread at lower redshift, likely due to the size growth of halos over time. 
% could make these same plots with scaled separation, then it would presumably show roughly the same spread for each redshift.

Likewise, fig.~\ref{fig:timevssep-high} presents the distribution for high-mass subhalos, where the trends between time until merger and separation are roughly the same. 
At each redshift, there is a positive correlation between time until merger and separation, but the spread on either of those decreases with increasing redshift.

These figures can be used to get approximate ranges of time until merger for a pair at a given separation and for a given redshift, and are useful for seeing the difference between the relationship between time until merger and separation for low-mass and high-mass pairs. \kc{what differences?} 
% 
\input{table-septime.tex}

\subsection{Redshift Dependence of Merger Timescales}\label{sec:results-timevredshift}
% The behavior of time until merger as a function of redshift for low-mass and high-mass pairs:
% 1 - the full sample 
% 2 - the physical and static kpc separation-selected sample
% 3 - the mass and redshift evolving scaled separation-selected sample
We study the merger timescale of low-mass and high-mass pairs as a function of redshift. 
In particular, we consider 3 different samples of the low- and high-mass systems. 
First, we will consider the orbit sample as a whole, and quantify the median merger timescale for all low- and high-mass pairs that are post-infall and pre-merger for $z=0-6$. 
We will then consider two different sets of separation criteria to create separation-selected subsamples, the first of which are physical separation criteria that are identically applied for both mass scales and all redshifts, and the second of which defines separation criteria that vary with mass and redshift.
% We ascertain the impact of the selection criteria on these subsamples of the orbit catalog ..... based on separation-selected pairs, using a set of separation criteria that we apply independently at each redshift.

Both separation-selected samples have a lower separation criteria of $10\,\kpc$, to limit the impact of subhalos becoming indistinguishable in the \texttt{SUBFIND} catalogs.
This lower separation criteria is also commonly applied to observationally-selected pairs in studies of merger fractions and merger rates (see Sec.~\ref{sec:discussion} for further detail).

Throughout, medians at each redshift are shown by solid pink and green lines, while the 1st and 3rd quartile spread on the median is shown by the shaded regions.

\subsubsection{Full sample}
We calculate the time until merger for all post-infall and pre-merger pairs at each redshift from $z=0-6$, and quantify the median and spread on the merger timescale as a function of redshift. 
We make no selection cuts based on pair separation, and thus include the full catalog of orbits at each redshift to calculate the merger timescale (once again, provided that they are post-infall and pre-merger at the given redshift). 
Essentially, we find the median time that any physically associated isolated pair will take to merge, provided it's a merging pair to begin with.

Figure~\ref{fig:timescales} presents the median, and 1st and 3rd quartile spread, of the time until merger for merging pairs from $z=0-6$. 
Low-mass merger timescales (green) and high-mass merger timescales (pink) are roughly equivalent at all redshifts. 
The median time until merger is $\sim0.5-0.7\,\Gyr$ at $z\sim6$, then rises to a peak of $2.3-2.4\,\Gyr$ at $z\sim0.6$, then decreases to 0 at $z=0$.

\kc{left off here}

% Pairs chosen at $z=1.5$ on average merge at $z=XX$/within XX Gyrs.

\subsubsection{Separation Criteria: Static \& Physical }
The first set of separation criteria will select only pairs at a given redshift that have separations greater than $10\,\kpc$ and less than $[50, 70, 100, 150, 200, \mbox{and }300]\kpc$. 
These separation criteria do not vary as a function of time or mass, and are thus applied equivalently to the low-mass and high-mass samples. 
This set of static physical separation criteria produces 6 subsamples of merging orbits that are post-infall.

Note that an orbit in the separation bin $10-50\,\kpc$ at $z=2$ will not necessarily be in that same bin at other redshifts. 
Rather, the separation criteria select pairs with separations in the given separation bin independently at each redshift.

% figure description 
The top panel of Fig.~\ref{fig:timescales-sep} shows the time until merger versus redshift for low-mass (green) and high-mass (pink) orbits that have separations less than the separation listed above each plot.  
% redshift evolution of each subsample
We first note that the evolution of the time until merger is approximately the same for each separation-selected subsample, with a peak time at redshift $z\sim0.5$ present for all subsamples, and a rapid decline of the median time until merger at $z=0-0.5$. 
These traits were also present in Fig.~\ref{fig:timescales} where the median time until merger was computed for the entire sample, without a selection-criterion.
As the selection-criteria gets larger and larger, the subsample will contain more and more of the full sample, eventually reconstructing the full sample plot with a large enough separation selection.

% low mass vs high mass
There is an offset between the median times for low-mass and high-mass pairs for each of the subsamples with maximum separations of $[50, 70, \mbox{and } 100]\,\kpc$. 
The difference in the median time until merger is up to $0.8\,\Gyr$ longer for low-mass pairs than high-mass pairs at the same redshift. 
The offset between the low-mass and high-mass merger times decreases for large enough physical separation criteria. 
In the rightmost top panel, the median merger times overlap almost identically for $z=0-0.5$ and $z>1.5$.

% impact of the separation criteria
The median time until merger for pairs selected in smaller separation bins (such as $10-50\,\kpc$ and $10-70\,\kpc$) is lower at all redshifts for both low-mass and high mass pairs. 
This follows directly from Sec.~\ref{sec:results-timevsep}, where we found that the time until merger is positively correlated with increasing separation for all pairs. 
Thus, including more and more larger-separation systems in our merger time calculation will tend to increase the median time until merger.

\subsubsection{Separation Criteria: Scaled Separation}
The second set of selection criteria will select pairs at a given redshift in 6 different scaled separation bins, where the scaled separation is written as a fraction of the virial radius of the pair's FoF group, $\Rvir$ (see Sec.~\ref{sec:methods} for more details). 
The scaled separation-selected pairs have separations greater than $10\,\kpc$ and less than $<[0.25, 0.5, 1, 1.5, 2, 2.5]\,\Rvir$.

A scaled separation criterion scales with both redshift and mass (since the virial radius scales with redshift and mass), and produces 6 subsamples of merging orbits that are post-infall.
As mentioned in \chambe{}, the median virial radius for low-mass FoF groups at $z=[0,1,2,3,4]$ is approximately $[134, 85, 59, 43, 33]\,\kpc$, and for high-mass FoF groups is approximately $[348, 206, 134, 97, 76]\,\kpc$.
Thus, choosing a scaled separation cut of $<1\,\Rvir$ at $z=1$ will select high-mass pairs with separations $\lesssim205\,\kpc$ and low-mass pairs with separations $\lesssim85\,\kpc$.

As in the previous section, the selection criteria are applied at each redshift independently. 
This means that the time until merger of a pair at a given redshift will only be included in the calculation of the median merger time if its separation is within the given separation bin at that redshift.

% figure description 
The bottom panel of Fig.~\ref{fig:timescales-sep} shows the median time until merger versus redshift for low-mass (green) and high-mass (pink) systems with scaled separations less than those listed above each plot. 
In the panel on the left, only pairs with a separation between $10\,\kpc-0.25\,\Rvir$ are utilized for the calculation. 
We find that the median time until merger at $z=1$ is XX..XX for each scaled separation cut respectively, for both low-mass and high-mass samples.

% redshift evolution of each subsample
Similar to the static physical separation-selected subsamples in the top panel, the median time until merger starts off small at $z>4$, then increases to a peak around $z\sim0.5$, before quickly decreasing to 0 at $z=0$. 
This behavior also mimics that of the full sample.

% impact of the separation criteria
Additionally, the median time until merger at a given redshift increases for separation criteria that include a larger fraction of the virial radius, and thus larger separation pairs. 
This again matches the behaviour of the static physical separation criteria shown in the top panel.

% low mass vs high mass
However, quite unlike the subsamples in the top panels, the scaled separation-selected subsamples yield nearly identical median merger timescales for low-mass and high-mass pairs at all redshifts, regardless of what fraction of the virial radius is selected.

@@

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_time_til_merger_full_fit.png}
    \caption{\kc{TODO: add LookbackTime vs. Z line, add Hubble time line} The median time until merger as a function of redshift for low-mass (green) and high-mass (pink) pairs. Shaded regions represent the 1st and 3rd quartile spread on the median. Pairs at each redshift are selected if they have experienced first infall. 
    % 
    The black dashed line shows the Hubble time, which bounds the median time until merger from $z=6$ to $z\sim0.5-0.75$.
    The dotted grey line shows the time to $z=0$ (the Lookback Time) as a function of redshift, which sets the upper bound for the time until merger that a merging pair can have. 
    The median time until merger is similar for low-mass and high-mass pairs, and rises from $z=6$ to a peak at $z\sim0.75$, then decreases to $z=0$.}
    \label{fig:timescales}
\end{figure*}

@@

\section{Discussion} \label{sec:discussion}

@@

\subsection{Results}

Why did we pick the separation cuts that we did? (motivated already in Chambe+ 2024, restate here about close pair definitions from lit.)

For a given static separation selection (particularly <50, <70, and <100), the merger timescales of low-mass and high-mass systems disagree by ~0.5\Gyr at almost all redshifts.

Scaled separation criteria are akin to picking the same fraction of the total volume of a FoF group, regardless of what redshift it is at, not what mass it has.

\subsection{What does it mean that the Time until Merge plot has the function form that it does?}
Upper bounds of the hubble time and the lookback time as functions of z.

@@

\section{Summary and Conclusions}
% what we did and how

% set of conclusions from results
Our main findings are as follows: 
\begin{itemize}
    \item for a given separation (such as 100kpc), high-mass pairs have shorter merger timescales than low-mass pairs. 
    \item the median time until merger for the full sample of low and high-mass pairs increases as $\sim(1+z)^{-1}$ from $z=6$ to $z\sim0.75$. it then decreases to 0 at low redshifts, with an upper bound set by the time remaining in the simulation before $z=0$. The decrease in the merger time at low redshifts is artificial and not representative of the true merger timescales that pairs at $z=0$ will have.
    \item separation criteria that include larger separations will increase the median time until merger.
    \item static physical separation-selected subsamples of low- and high-mass pairs have different median merger timescales.
    \item scaled separation-selected subsamples yield nearly identical median merger timescales for low-mass and high-mass pairs at all redshifts. 
\end{itemize}

% set of conclusions for discussion
- We find that the merger timescale of low-mass or high-mass merging pairs is sensitive to the selection criteria that are used to define the pair sample.
- We recommend that future studies that seek to compare or compute pair fractions, merger fractions, merger timescales, or merger rates, define selected pair samples using separation criteria that evolve with redshift and mass.

% ** 
% \appendix
% adfadfa
